Bibat: Batteries-include Bayesian Analysis Template

### 2024-01-30 paper/main.tex
Update paper

\documentclass[sigconf, review]{acmart}

\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother

%%
%% This is file `sample-manuscript.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `manuscript')
%%
%% IMPORTANT NOTICE:
%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-manuscript.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass command.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}

\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}

%% PANDOC PREAMBLE BEGINS

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\definecolor{mypink}{RGB}{219, 48, 122}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
%% PANDOC PREAMBLE ENDS

\setlength{\parindent}{10pt}
\setlength{\parskip}{0pt}

\hypersetup{
  pdftitle={Bibat: Batteries-include Bayesian Analysis Template},
  pdfauthor={Teddy Groves},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={red},
  pdfcreator={LaTeX via pandoc, via quarto}}

%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2024}
\acmYear{2024}
\acmDOI{}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[KDD]{KDD}{Aug 25--29, 2024}{Barcelona, Esp}
\acmPrice{}
\acmISBN{}

%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Bibat: Batteries-include Bayesian Analysis Template}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.

  \author{Teddy Groves}
  \orcid{0000-0002-7109-3270}
            \affiliation{%
                  \institution{Danish Technical University}
                                  \city{Kongens Lyngby}
                                  \country{Denmark}
                      }

%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
%\renewcommand{\shortauthors}{Trovato et al.}
%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
Bayesian statistical workflow offers a powerful way to learn from data,
but software software projects that implement complex Bayesian workflows
in practice are unusual, partly due to the difficulty of orchestrating
Bayesian statistical software. Bibat addresses this challenge by
providing a full-featured, scalable Bayesian statistical analysis
project using an interactive template. Bibat is available on the Python
Package index, documented at \url{https://bibat.readthedocs.io/} and
developed at \url{https://github.com/teddygroves/bibat/}. This paper
explains the motivation for bibat, briefly describes intended usage,
discusses key design choices, and reviews several examples of bibat's
use in scientific applications.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Python, Bayesian workflow, Methodology}

\begin{teaserfigure}
  \includegraphics[width=\textwidth]{img/workflow.png}
  \caption{Schematic representation of a Bayesian workflow implemented
using bibat. The author inspects their analysis's results, edits files
corresponding to the boxes on the left, runs the command
\texttt{make\ analysis}, then repeats. Note that this workflow is
modular, accommodates plurality and results in a final analysis that can
be fully reproduced using a single command.}
  \Description{DESCRIPTION}
\end{teaserfigure}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\setlength{\parskip}{-0.1pt}

@@

The idea that Bayesian statistical analysis comprises not just
inference, but also specific approaches to related activities like data
preparation, model design, diagnosis, debugging and criticism, dates at
least to \citet{boxBayesianInferenceStatistical1992}. There is
increasing scholarly recognition of the need for a holistic view of
``Bayesian workflow''
\citep{gelmanBayesianWorkflow2020, grinsztajnBayesianWorkflowDisease2021, gabryVisualizationBayesianWorkflow2019}.
Relatedly, software tools now exist that address most individual aspects
of a Bayesian workflow: see \citet{strumbeljPresentFutureSoftware} for a
review of the state of the art.

@@

Bibat can be installed on the Windows, linux or macos command line on by
running the command \texttt{pip\ install\ bibat} and then used by
running the command \texttt{bibat}. This command triggers an interactive
form which prompts the user to select a range of customisation options.
Bibat then creates a new directory containing code that implements an
example analysis, with customisations reflecting the user's choices.
This analysis works immediately, and can be reproduced with the single
command \texttt{make\ analysis} without the need for any further action
by the user: in this sense bibat comes with ``batteries included''.

@@

@@

Bibat projects are is written in modern Python and uses pydantic
\citet{pydanticdevelopersPydantic2022} and pandera
\citep{niels_bantilan-proc-scipy-2020} for data modelling, Stan
\citep{carpenterStanProbabilisticProgramming2017} for statistical
inference, cmdstanpy \citep{standevelopmentteamCmdStanPy2022} for
python-Stan interface, arviz \citep{kumarArviZUnifiedLibrary2019} for
storing and analyzing inferences and sphinx
\citep{georgbrandlandthesphinxteamSphinx2022} and quarto
\citep{Allaire_Quarto_2022} for documentation.

Bibat itself is also written in modern Python and uses the popular tools
cookiecutter \citep{greenfeldCookiecutter2021}, pydantic and click
\citep{clickdevelopersClickPythonComposable2022}.

Bibat's makefile detects the current operating system and attempts to
install cmdstan appropriately if necessary. This functionality addresses
a common issue where researchers find it difficult to install Stan,
especially on Windows.

@@

Bibat is part of the PyOpenSci ecosystem, allowing for community help
with maintenance as well as peer review for code and documentation
quality, usability and accessibility. The PyOpenSci peer review for
bibat can be found here:
\url{https://github.com/pyOpenSci/software-submission/issues/83}

@@

As discussed in \citet{gelmanBayesianWorkflow2020}, Bayesian workflows
are complicated, featuring plurality, cyclicity and complexity at many
levels. Bibat accommodates this by separating non-interacting analysis
components into modules and by serialising data to files wherever
possible. Prepared datasets, statistical models, inference
configurations, inference results, plots and analyses all have file
representations. Fitting modes, data manipulations and data models are
modularised in code through the use of appropriately structured data
classes and functions.

@@

\section{Post launch use}\label{post-launch-use}

Bibat has been used for several real statistical analyses.

\citet{grovesteddyDgfreg2023} used bibat to compare a Bayesian and two
non-Bayesian approaches to modelling a biochemical thermodynamics
dataset. Bibat facilitated this analysis even though it was not very
large by providing the fitting mode abstraction, which was very useful
for comparing the different methods.

In \citet{grovesBaseball2022}, Bibat was used to implement a sports
analysis involving two datasets, two models and four inferences,
demonstrating that the generalised Pareto distribution can be used to
describe hitting ability in baseball. This analysis is now included in
bibat as an illustration, along with an accompanying tutorial.

@@

\includegraphics[width=0.5\textwidth,height=\textheight]{img/baseball.png}

@@

In \citet{grovesteddySphincter2024}, bibat was used to implement an
analysis of cerebrovascular data from mice, involving two raw datasets,
6 prepared datasets, 15 models and 15 inferences. In this case bibat's
modular design made it straightforward to iteratively add data
transformations and models while maintaining reproducibility.

@@

Bibat has a growing user community, with 16 GitHub stars at the time of
writing.

Other than bibat, there is currently no interactive template that
specifically targets Bayesian workflow projects. There are some
templates that arguably encompass Bayesian workflow as a special case of
data analysis project, such as cookiecutter-data-science
\citep{drivendataCookiecutterdatascience2022}, but these are of limited
use compared with a specialised template due to the many specificities
of Bayesian workflow.

There is some software that addresses the general task of facilitating
Bayesian workflow, but using a different approach from bibat's. For
example, bambi \citep{capretto2020} and brms
\citep{burknerBrmsPackageBayesian2017} aim to make implementing Bayesian
workflows easier by providing ergonomic ways to specify and fit Bayesian
regression models to tabular datasets. Bibat is complementary with these
packages, as it targets use cases that they do not support, such as
analyses where complex datasets or custom models might be required.

\section{Statements and Declarations}\label{statements-and-declarations}

The authors have no financial or non-financial interests that are
directly or indirectly related to the work submitted for publication.

\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography.bib}

%% begin pandoc before-bib
%% end pandoc before-bib
%% begin pandoc biblio
%% end pandoc biblio
%% begin pandoc include-after
%% end pandoc include-after
%% begin pandoc after-body
%% end pandoc after-body

\end{document}
\endinput
%%
%% End of file `sample-manuscript.tex'.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Bibat: Batteries-include Bayesian Analysis Template},
  pdfauthor={Teddy Groves},
  pdfkeywords={Bayesian workflow, Methodology, Software},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Bibat: Batteries-include Bayesian Analysis Template}
\author{Teddy Groves}
\date{}

\begin{document}
\maketitle
\begin{abstract}
Bayesian statistical workflow offers a powerful way to learn from data,
but software software projects that implement complex Bayesian workflows
in practice are unusual, partly due to the difficulty of orchestrating
Bayesian statistical software. Bibat addresses this challenge by
providing a full-featured, scalable Bayesian statistical analysis
project using an interactive template. Bibat is available on the Python
Package index, documented at \url{https://bibat.readthedocs.io/} and
developed at \url{https://github.com/teddygroves/bibat/}. This paper
explains the motivation for bibat, describes intended usage, discusses
bibat's design goals and how key challenges that were overcome in the
course of development, compares bibat with similar software, highlights
several examples of bibat's use in science and provides links to
community resources associated with bibat.
\end{abstract}

@@

The idea that Bayesian statistical analysis comprises not just
inference, but also specific approaches to related activities like data
preparation, model design, diagnosis, debugging and criticism, dates at
least to Box and Tiao (1992). There is increasing scholarly recognition
of the need for a holistic view of ``Bayesian workflow'' (Gelman et al.
2020; Grinsztajn et al. 2021; Gabry et al. 2019). Relatedly, software
tools now exist that address most individual aspects of a Bayesian
workflow: see Štrumbelj et al. (2024) for a review of the state of the
art.

@@

Bibat can be installed on the Windows, Linux or macOS command line on by
running the command \texttt{pip\ install\ bibat} and then used by
running the command \texttt{bibat}. This command triggers an interactive
form which prompts the user to select a range of customisation options.
Bibat then creates a new directory containing code that implements an
example analysis, with customisations reflecting the user's choices.
This analysis works immediately, and can be reproduced with the single
command \texttt{make\ analysis} without the need for any further action
by the user: in this sense bibat comes with batteries included.

Figure~\ref{fig-workflow} illustrates the components of a bibat-based
Bayesian workflow and shows how it proceeds: the project team edits the
code components, then runs \texttt{make\ analysis} triggering creation
of the result components. After inspecting these they repeat the
process, leading to a cycle that ultimately results in a complete,
easily reproducible analysis.

\includegraphics[width=1\textwidth,height=\textheight]{img/workflow.png}

\caption{\label{fig-workflow}Schematic representation of a Bayesian
workflow implemented using bibat. The author inspects their analysis's
results, edits code corresponding to the boxes on the left, runs the
command \texttt{make\ analysis}, then repeats. The diagram illustrates
several key features of bibat: inference components are modular and
plural, the overall workflow is iterative and cyclical and the whole
analysis can be executed with a single command.}

@@

In particular, this target attempts to install cmdstan if necessary,
using a recipe tailored to the host operating system. This functionality
addresses a common issue where researchers find it difficult to install
Stan, especially on Windows.

@@

Bibat projects are is written in modern Python and uses pydantic
Pydantic developers (2022) and pandera (Niels Bantilan 2020) for data
modelling, Stan (Carpenter et al. 2017) for statistical inference,
cmdstanpy (Stan Development Team 2022) for python-Stan interface, arviz
(Kumar et al. 2019) for storing and analyzing inferences and sphinx
(Georg Brandl and the Sphinx team 2022) and quarto (Allaire et al. 2022)
for documentation.

Bibat itself is also written in modern Python and uses the popular tools
cookiecutter (Greenfeld et al. 2021), pydantic and click (Click
Developers 2022).

@@

@@

As discussed in Gelman et al. (2020), Bayesian workflows are
complicated, featuring plurality, cyclicity and complexity at many
levels. Bibat accommodates this by separating non-interacting analysis
components into modules and by serialising data to files wherever
possible. Prepared datasets, statistical models, inference
configurations, inference results, plots and analyses all have file
representations. Fitting modes, data manipulations and data models are
modularised in code through the use of appropriately structured data
classes and functions.

@@

Other than bibat, there is currently no interactive template that
specifically targets Bayesian workflow projects. There are some
templates that arguably encompass Bayesian workflow as a special case of
data analysis project, such as cookiecutter-data-science (Driven Data
2022), but these are of limited use compared with a specialised template
due to the many specificities of Bayesian workflow.

There is some software that addresses the general task of facilitating
Bayesian workflow, but using a different approach from bibat's. For
example, bambi (Capretto et al. 2020) and brms (Bürkner 2017) aim to
make implementing Bayesian workflows easier by providing ergonomic ways
to specify and fit Bayesian regression models to tabular datasets. Bibat
is complementary with these packages, as it targets use cases that they
do not support, such as analyses where complex datasets or custom models
might be required.

\section{Case studies}\label{case-studies}

The following cases illustrate how bibat has been used in practice to
facilitate Bayesian workflow projects.

Groves and Jooste (2023) used bibat to compare a Bayesian and two
non-Bayesian approaches to modelling a biochemical thermodynamics
dataset. Bibat facilitated this analysis even though it was not very
large---the final analysis contained one dataset, three models and three
inferences---because the fitting mode abstraction allowed for
straightforward comparison of the different methods. Additionally, bibat
made it easier to iteratively investigate and discard models that did
not form part of the final analysis.

In Groves (2022), Bibat was used to implement a sports analysis
involving two datasets, two models and four inferences, demonstrating
that the generalised Pareto distribution can be used to describe hitting
ability in baseball. This analysis is now included in bibat as an
illustration, along with an accompanying tutorial. An illustrative
graphic from this analysis is shown in Figure~\ref{fig-baseball}.

In this case bibat was useful because of its ability to implement
arbitrary statistical models, as latent generalised Pareto distributions
are not supported by generic regression packages like brms and bambi.
Further, bibat's modular design made it easier to implement this
medium-sized analysis with two datasets, two models and six inferences.

@@

\includegraphics[width=1\textwidth,height=\textheight]{img/baseball.png}

@@

In Groves (2024), bibat was used to implement a large analysis of
cerebrovascular data from mice, involving two raw datasets, 6 prepared
datasets, 15 models and 15 inferences. Again it was not possible to
implement the analysis using formula-based packages due to the need for
custom modelling approaches to take into account expert knowledge about
both the target system and the measurement apparatus. Due to the number
of datasets and models that needed to be considered, implementing this
analysis without using bibat or a similar structured approach would have
required a combination of more time and/ or compromises in software
quality.

@@

\section{Community}\label{community}

Bibat is developed in public and encourages community contribution.
Please see the contributing page
\url{https://github.com/teddygroves/bibat/blob/main/CONTRIBUTING.rst}
and code of conduct
\url{https://github.com/teddygroves/bibat/blob/main/CODE_OF_CONDUCT.md}
if you would like to help develop bibat.

Bibat has a growing user community, with 16 GitHub stars at the time of
writing, and is affiliated with cmdstanpy through a link on its
documentation website. Bibat is also affiliated with the Python
scientific software community PyOpenSci, allowing for help with
maintenance as well as peer review for code and documentation quality,
usability and accessibility. The PyOpenSci peer review for bibat can be
found here:
\url{https://github.com/pyOpenSci/software-submission/issues/83}.

\section{Author contributions}\label{author-contributions}

\section*{Acknowledgements}\label{acknowledgements}
\addcontentsline{toc}{section}{Acknowledgements}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-Allaire_Quarto_2022}
Allaire, J. J., Charles Teague, Carlos Scheidegger, Yihui Xie, and
Christophe Dervieux. 2022. {``Quarto.''}
\url{https://doi.org/10.5281/zenodo.5960048}.

\bibitem[\citeproctext]{ref-boxBayesianInferenceStatistical1992}
Box, George E. P., and George C. Tiao. 1992. \emph{Bayesian Inference in
Statistical Analysis}. Wiley classics library ed. A {Wiley-Interscience}
Publication. {New York}: {Wiley}.

\bibitem[\citeproctext]{ref-burknerBrmsPackageBayesian2017}
Bürkner, Paul-Christian. 2017. {``Brms: {An R} Package for {Bayesian}
Multilevel Models Using {Stan}.''} \emph{Journal of Statistical
Software} 80 (1): 1--28. \url{https://doi.org/10.18637/jss.v080.i01}.

\bibitem[\citeproctext]{ref-capretto2020}
Capretto, Tomás, Camen Piho, Ravin Kumar, Jacob Westfall, Tal Yarkoni,
and Osvaldo A. Martin. 2020. {``Bambi: {A} Simple Interface for Fitting
{Bayesian} Linear Models in {Python}.''}
\url{https://arxiv.org/abs/2012.10754}.

\bibitem[\citeproctext]{ref-carpenterStanProbabilisticProgramming2017}
Carpenter, Bob, Andrew Gelman, Matthew D. Hoffman, Daniel Lee, Ben
Goodrich, Michael Betancourt, Marcus Brubaker, Jiqiang Guo, Peter Li,
and Allen Riddell. 2017. {``Stan: {A Probabilistic Programming
Language}.''} \emph{Journal of Statistical Software} 76 (1): 1--32.
\url{https://doi.org/10.18637/jss.v076.i01}.

\bibitem[\citeproctext]{ref-clickdevelopersClickPythonComposable2022}
Click Developers. 2022. {``Click: {Python} Composable Command Line
Interface Toolkit.''} Pallets. \url{https://pypi.org/project/click/}.

\bibitem[\citeproctext]{ref-drivendataCookiecutterdatascience2022}
Driven Data. 2022. {``Cookiecutter-Data-Science.''}
\url{https://github.com/drivendata/cookiecutter-data-science/}.

\bibitem[\citeproctext]{ref-gabryVisualizationBayesianWorkflow2019}
Gabry, Jonah, Daniel Simpson, Aki Vehtari, Michael Betancourt, and
Andrew Gelman. 2019. {``Visualization in {Bayesian Workflow}.''}
\emph{Journal of the Royal Statistical Society Series A: Statistics in
Society} 182 (2): 389--402. \url{https://doi.org/10.1111/rssa.12378}.

\bibitem[\citeproctext]{ref-gelmanBayesianWorkflow2020}
Gelman, Andrew, Aki Vehtari, Daniel Simpson, Charles C. Margossian, Bob
Carpenter, Yuling Yao, Lauren Kennedy, Jonah Gabry, Paul-Christian
Bürkner, and Martin Modrák. 2020. {``Bayesian {Workflow}.''}
\emph{arXiv:2011.01808 {[}Stat{]}}, November.
\url{http://arxiv.org/abs/2011.01808}.

\bibitem[\citeproctext]{ref-georgbrandlandthesphinxteamSphinx2022}
Georg Brandl and the Sphinx team. 2022. {``Sphinx.''}
\url{https://www.sphinx-doc.org/}.

\bibitem[\citeproctext]{ref-greenfeldCookiecutter2021}
Greenfeld, Audrey Roy, Dainiel Roy Greenfeld, Raphael Pierzina, et al.
2021. {``Cookiecutter.''} \url{https://pypi.org/project/cookiecutter/}.

\bibitem[\citeproctext]{ref-grinsztajnBayesianWorkflowDisease2021}
Grinsztajn, Léo, Elizaveta Semenova, Charles C. Margossian, and Julien
Riou. 2021. {``Bayesian Workflow for Disease Transmission Modeling in
{Stan}.''} \emph{Statistics in Medicine} 40 (27): 6209--34.
\url{https://doi.org/10.1002/sim.9164}.

\bibitem[\citeproctext]{ref-grovesBaseball2022}
Groves, Teddy. 2022. {``Baseball.''}
\url{https://github.com/teddygroves/baseball}.

\bibitem[\citeproctext]{ref-grovesteddySphincter2024}
---------. 2024. {``Sphincter.''}
\url{https://github.com/teddygroves/sphincter}.

\bibitem[\citeproctext]{ref-grovesteddyDgfreg2023}
Groves, Teddy, and Jason Jooste. 2023. {``Dgfreg.''} DTU Biosustain.
\url{https://github.com/biosustain/dgfreg}.

\bibitem[\citeproctext]{ref-kumarArviZUnifiedLibrary2019}
Kumar, Ravin, Colin Carroll, Ari Hartikainen, and Osvaldo Martin. 2019.
{``{ArviZ} a Unified Library for Exploratory Analysis of {Bayesian}
Models in {Python}.''} \emph{Journal of Open Source Software} 4 (33):
1143. \url{https://doi.org/10.21105/joss.01143}.

\bibitem[\citeproctext]{ref-niels_bantilan-proc-scipy-2020}
Niels Bantilan. 2020. {``Pandera: {Statistical Data Validation} of
{Pandas Dataframes}.''} In \emph{Proceedings of the 19th {Python} in
{Science Conference}}, edited by Meghann Agarwal, Chris Calloway, Dillon
Niederhut, and David Shupe, 116--24.
\url{https://doi.org/10.25080/Majora-342d178e-010}.

\bibitem[\citeproctext]{ref-pydanticdevelopersPydantic2022}
Pydantic developers. 2022. {``Pydantic.''}
\url{https://pypi.org/project/pydantic/}.

\bibitem[\citeproctext]{ref-standevelopmentteamCmdStanPy2022}
Stan Development Team. 2022. {``{CmdStanPy}.''}
\url{https://github.com/stan-dev/cmdstanpy}.

\bibitem[\citeproctext]{ref-strumbeljPresentFutureSoftware}
Štrumbelj, Erik, Alexandre Bouchard-Côté, Jukka Corander, Andrew Gelman,
Håvard Rue, Lawrence Murray, Henri Pesonen, Martyn Plummer, and Aki
Vehtari. 2024. {``Past, {Present}, and {Future} of {Software} for
{Bayesian Inference}.''}

\end{CSLReferences}

\end{document}
